---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BayesianDFM

<!-- badges: start -->
<!-- badges: end -->

The BayesianDFM package provides relevant functions for running a Bayesian
Dynamic Factor Model.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}

devtools::install_github("h4sci/packagr")
```


## Functions

The following functions are contained in the package:

  * BVAR_Jeff
  * compFstate
  * drawLam
  * drawSig
  * get_IC
  * get_nowcast
  * multimoveGibbs
  * prepare_data
  * run_model

For further information about the functions, check out the function
descriptions.



## Example

This is a basic example which shows you how to solve a common problem:

Load and prepare the input data
```{r, eval = FALSE}

load("data_UK.Rda")

target <- c("UKGDPM.YQ") # Define target variable

# De-mean & standardize data
Xmat <- prepare_data(flows = data$flows,
                     stocks = data$stocks,
                     inventory,
                     target = target)

yt <- as.matrix(t(Xmat))

```

Make a in-sample evaluation of the optimal number of factors

```{r, eval = FALSE}
IC <- get_IC(Xmat) # Check information criteria

```

Define the parameters for the model

```{r, eval = FALSE}
k <- 2 # number of states (number of factors)

q <- 1 # lag length for state equation
m <- k*q

n <- dim(yt)[1] # Number of variables
Tt <- dim(yt)[2] # Number of high-frequency periods
Ttq <- Tt-q
const <- 0 # choose constant in the state equation (we choose no constant)

```

Having all the parameters, run the model

```{r, eval = FALSE}
out <- run_model(yt,k,q,m,n,Tt,Ttq,const,target)

```

